---
title: "Reproducible Research P1"
author: "Phil Burdi (pburdi@gmail.com)"
date: "Sunday, January 18, 2015"
output: html_document
---


Load up the libraries needed for this assignment.
```{r, echo = FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)

```

Download the data file and after we downloaded the file, we unzipped manual using winwar. (download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "c5p1.zip"))

Read the data file and setup the environment.  

```{r}
ACT1 <- read.csv("activity.csv", sep = ",", header = TRUE)
ACT2 <- na.omit(ACT1)
ACT2 <- group_by(ACT2, date)
summary(ACT2)
```

We are going to summarize to get the median and mean per day.  For question "What is mean total number of steps taken per day?"

```{r}
ACT3 <- summarize(ACT2, total_steps = sum(steps), mean_steps = mean(steps), median_steps = median(steps))
summary(ACT3$total_steps)
```

Now produce the histogram to show the total steps during all the intervals.

```{r}
hist(ACT3$total_steps, xlab="5 Minute Interval", main="Histogram of Total Steps")
```

To answer the question "What is the average daily activity pattern?" we applied the mean to the steps and to the interval`.

```{r}
ACT4 <- tapply(ACT2$steps, ACT2$interval, mean)
```

Now plot the result of the average number of steps taken.

```{r}
plot(row.names(ACT4), ACT4, type = "l", xlab = "5-min a day interval", ylab = "Average Steps ", main = "Average Number Of Steps Taken", col = "red")
```

Now we input the missing values and see how it skews the results - we check the max interval and the number of rows

```{r}
Max_Interval <- which.max(ACT4)
names(Max_Interval)
na_cases <- ACT1[!complete.cases(ACT1),]
nrow(na_cases)
```

Lets assign some data to the NAs

```{r}
ACT1_1 <- ACT1
ACT1_1[is.na(ACT1_1)] <- 1
```

Plot the result and check against the previous histogram

```{r}
ACT2_1 <- group_by(ACT1_1, date)
ACT3_1 <- summarize(ACT2_1, total_steps = sum(steps), mean_steps = mean(steps), median_steps = median(steps))
summary(ACT3_1$total_steps)
hist(ACT3_1$total_steps, xlab="5 Minute Interval", main="Total Steps with Data Added to NAs")
```

Now that we have seen the results above and plotted, the answer to this question "Do these values differ from the estimates from the first part of the assignment?" YES - but not by much on the mean and median, however the 0-5000 interval was significantly changed.

Also, the question "What is the impact of imputing missing data on the estimates of the
total daily number of steps?" It skews the lower end of the frequency.

-----------------------------------------------------------------------

Looking at the final question "Are there differences in activity patterns between weekdays and weekends?" we will setup the data-sets to identify the weekday and weekend patterns

```{r}
ACT2_1$date <- as.Date(ACT2_1$date, "%Y-%m-%d")
ACT2_1$day <- weekdays(ACT2_1$date)
ACT2_1$day_type <- c("weekday")
for (w in 1:nrow(ACT2_1)){
   if (ACT2_1$day[w] == "Saturday" || ACT2_1$day[w] == "Sunday"){
     ACT2_1$day_type[w] <- "weekend"
   }
}
ACT2_2 <- aggregate(steps ~ interval+day_type, ACT2_1, mean)
```

Now lets examine the data-set in a plot.  

```{r}
qplot(interval, steps, data=ACT2_2, geom=c("line"), xlab="5 Minute Interval", ylab="Number of Steps", main="") + facet_wrap(~ day_type, ncol=1)
```

As shown the data for the weekday plot indicates a high frequency of activity in the morning and tapers off significantly around 10 AM - the weekend plot indicates a fairly
consistent pattern from morning to mid-afternoon.

Thank you for reading - end of line...